---
title: "Transverse Ball-Bat Collision"
output: html_document
date: "2022-08-22"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r}
#Import packages
library(ggplot2)
library(dplyr)
library(ggrepel)
library(metR)
```

## Hawkeye Data
```{r}
hawk_data <- read.csv("C:/Users/12244/Baseball_2023/Alan_bat_data.csv")
hawk_data_inplay <- hawk_data %>%
  filter(hit_into_play_group != "")

hawk_data_inplay_player <- hawk_data_inplay %>%
  group_by(batter_name) %>%
  dplyr::summarize(mean_attack_angle = mean(sample_vert_attack_angle),
                   median_attack_angle = median(sample_vert_attack_angle),
                   count = n()) %>%
  filter(count > 60)
```

# 1. Data Preparation
## Constants
```{r}
#Constants:
pitch_speed = 84.8 #mph (84.8, 93.9, 55)
descent_angle = 8.0 #degrees (8.0, 5, 5)
bat_speed = 73.0 #mph (73.0, 80.0, 200)
attack_angle = 12.0 #degrees
pitch_spin = -2000 #rpm (-2000, -1500)

mball = 5.125 #oz
mbat = 31.494 #oz
rball = 1.452 #in
rbat = 1.267 #in

gamma = descent_angle * pi / 180 #rad
psi = attack_angle * pi / 180 #rad

mu = 0.5500
alpha = 0.400
mbateff_x = 7.817 #oz
recoilx = 2 / 7 * mball / mbateff_x

v0ball = pitch_speed * 5280 * 12 / 3600 #ft/s
v0bat = bat_speed * 5280 * 12 / 3600 #ft/s
mbateff_y = 23.172 #oz

recoily = mball / mbateff_y

ex = 0.300
eT = (ex - recoilx) / (1 + recoilx)
omegai = pitch_spin * pi / 30 #rad/s 

```

## Function for Generating Data from Offsets
```{r}
#Function for obtaining data with offset as input
generate_data_from_offset <- function(offset) {
  theta = asin(offset / (rball + rbat)) #rad
  centerline_angle = theta * 180 / pi #degree
  D = (rball + rbat) * sin(theta - psi) #inch
  v0ball_perp = v0ball * cos(theta - gamma) #ft/s
  v0bat_perp = v0bat * cos(theta - psi) #ft/s
  ey = 0.58 - (v0ball_perp + v0bat_perp - 1055) / 14094
  eA = (ey - recoily) / (1 + recoily)
  v0ball_T = v0ball * sin(theta - gamma) #ft/s
  v0bat_T = - v0bat * sin(theta - psi) #ft/s
  
  mulim = 0.55
    
  vfball_perp = v0ball_perp * eA + v0bat_perp * (1 + eA) #ft/s
  vfball_T = v0ball_T * (1 - alpha * eT) / (1 + alpha) + rball * omegai * alpha * (1 + eT) / (1 + alpha) + v0bat_T * alpha * (1 + eT) / (1 + alpha) #ft/s
  
  angle = atan(vfball_T / vfball_perp) + theta #rad
  
  launch_angle = angle * 180 / pi # degrees
    
  omegaf = omegai + (v0ball_T - vfball_T) / (alpha * rball) #rad / s
  
  exit_speed = sqrt(vfball_perp^2 + vfball_T^2) / (5280 * 12 / 3600) #mph
  
  exit_spin = omegaf * 30 / pi #rpm
  
  mulim = abs(v0ball_T - vfball_T) / (v0ball_perp + vfball_perp)
  
  df <- data.frame(offset = offset,
                   centerline_angle = centerline_angle,
                   D = D,
                   exit_speed = exit_speed,
                   exit_spin = exit_spin,
                   launch_angle = launch_angle,
                   v0ball_perp = v0ball_perp,
                   v0bat_perp = v0bat_perp,
                   ey = ey,
                   eA = eA,
                   v0ball_T = v0ball_T,
                   v0bat_T = v0bat_T,
                   omegaf = omegaf,
                   vfball_perp = vfball_perp,
                   vfball_T = vfball_T,
                   angle = angle,
                   mulim = mulim)
                   
  return(df)
}
```

## Generating Data from Offsets
```{r}
#Make new column with column names of data
columns <- c('offset', 'centerline_angle', 'D', 'exit_speed', 'exit_spin', 'launch_angle', 'v0ball_perp', 'v0bat_perp', 'ey', 'eA', 'v0ball_T', 'v0bat_T', 'omegaf', 'vfball_perp', 'vfball_T', 'angle', 'mulim')

df <- data.frame(matrix(nrow = 0, ncol = length(columns)))

colnames(df) = columns

#Offset values
offset_values = seq(0, 2.60, by = 0.05)

#Obtain data for every offset values
for (value in offset_values) {
  new_df <- generate_data_from_offset(value)
  df <- rbind(df, new_df)
}
```

## Function for Generating Data from CA & AA
```{r}
#Function for obtaining data with center line angle and attack angle as parameters
generate_data_from_ca_aa <- function(CA, AA) {
  psi = AA * pi / 180 #rad
  theta = CA / 180 * pi #rad
  offset = sin(theta) * (rball + rbat) #inch
  D = (rball + rbat) * sin(theta - psi) #inch
  v0ball_perp = v0ball * cos(theta - gamma) #ft/s
  v0bat_perp = v0bat * cos(theta - psi) #ft/s
  ey = 0.58 - (v0ball_perp + v0bat_perp - 1055) / 14094
  eA = (ey - recoily) / (1 + recoily)
  v0ball_T = v0ball * sin(theta - gamma) #ft/s
  v0bat_T = - v0bat * sin(theta - psi) #ft/s
  
  mulim = 0.55
    
  vfball_perp = v0ball_perp * eA + v0bat_perp * (1 + eA) #ft/s
  vfball_T = v0ball_T * (1 - alpha * eT) / (1 + alpha) + rball * omegai * alpha * (1 + eT) / (1 + alpha) + v0bat_T * alpha * (1 + eT) / (1 + alpha) #ft/s
  
  angle = atan(vfball_T / vfball_perp) + theta #rad
  
  launch_angle = angle * 180 / pi # degrees
    
  omegaf = omegai + (v0ball_T - vfball_T) / (alpha * rball) #rad / s
  
  exit_speed = sqrt(vfball_perp^2 + vfball_T^2) / (5280 * 12 / 3600) #mph
  
  exit_spin = omegaf * 30 / pi #rpm
  
  mulim = abs(v0ball_T - vfball_T) / (v0ball_perp + vfball_perp)
  
  df <- data.frame(AA = AA,
                   CA = CA,
                   offset = offset,
                   D = D,
                   exit_speed = exit_speed,
                   exit_spin = exit_spin,
                   launch_angle = launch_angle,
                   v0ball_perp = v0ball_perp,
                   v0bat_perp = v0bat_perp,
                   ey = ey,
                   eA = eA,
                   v0ball_T = v0ball_T,
                   v0bat_T = v0bat_T,
                   omegaf = omegaf,
                   vfball_perp = vfball_perp,
                   vfball_T = vfball_T,
                   angle = angle,
                   mulim = mulim)
  return(df)
}
```

## Generating Data from CA & AA
```{r}
#Make new column with column names of data
columns_AA_CA <- c('AA', 'CA', 'offset', 'D', 'exit_speed', 'exit_spin', 'launch_angle', 'v0ball_perp', 'v0bat_perp', 'ey', 'eA', 'v0ball_T', 'v0bat_T', 'omegaf', 'vfball_perp', 'vfball_T', 'angle', 'mulim')

data_AA_CA <- data.frame(matrix(nrow = 0, ncol = length(columns_AA_CA)))

colnames(data_AA_CA) = columns_AA_CA

#Offset values
offset_AA = seq(-10, 50, by = 0.5)
offset_CA = seq(-10, 40, by = 0.5)

#Obtain data for every offset values
for (AA in offset_AA) {
  for (CA in offset_CA) {
    new_data_AA_CA <- generate_data_from_ca_aa(CA, AA)
    data_AA_CA <- rbind(data_AA_CA, new_data_AA_CA)
  }
}

#Add scaled exit velocity to data
data_AA_CA <- data_AA_CA %>%
  mutate(scaled_ev = exit_speed / max(data_AA_CA$exit_speed),
         scaled_ev = round(scaled_ev, 4))

#Add scaled exit velocity per AA to data
data_AA_CA <- data_AA_CA %>%
  group_by(AA) %>%
  mutate(max_exit_speed = max(exit_speed)) %>%
  ungroup() %>%
  mutate(scaled_ev_by_AA = exit_speed / max_exit_speed)
  
# Data with only max ev for each AA
data_AA_CA_max_ev <- data_AA_CA %>%
  group_by(AA) %>%
  filter(exit_speed == max(exit_speed))
```

## Function for Filtering Launch Angle
```{r}
#Function to filter launch angle
filter_data_LA <- function(LA) {
  data_LA <- data_AA_CA %>%
    filter(round(launch_angle) == LA)
  return(data_LA)
}

```

# 2. Plotting Data
## Plot Data
```{r, message = FALSE}
ggplot(data_AA_CA, aes(CA, AA, z = exit_speed)) + geom_contour(binwidth = 1) +
   geom_text_contour() + geom_abline(CA = AA) + 
  geom_smooth(data = filter_data_LA(0), se = FALSE, linetype = "dashed", color = "red") + 
  geom_smooth(data = filter_data_LA(10), se = FALSE, linetype = "dashed", color = "red") + 
  geom_smooth(data = filter_data_LA(20), se = FALSE, linetype = "dashed", color = "red") + 
  geom_smooth(data = filter_data_LA(30), se = FALSE, linetype = "dashed", color = "red") + 
  geom_smooth(data = filter_data_LA(40), se = FALSE, linetype = "dashed", color = "red") + 
  geom_smooth(data = filter_data_LA(50), se = FALSE, linetype = "dashed", color = "red") + 
  geom_text_repel(data = filter_data_LA(0)[nrow(filter_data_LA(0)), ], aes(label = paste("LA: ", round(launch_angle))), color = "red", position = "identity") + 
  geom_text_repel(data = filter_data_LA(10)[nrow(filter_data_LA(10)), ], aes(label = round(launch_angle)), color = "red", position = "identity") +
  geom_text_repel(data = filter_data_LA(20)[nrow(filter_data_LA(20)), ], aes(label = round(launch_angle)), color = "red", position = "identity") + 
  geom_text_repel(data = filter_data_LA(30)[nrow(filter_data_LA(30)), ], aes(label = round(launch_angle)), color = "red", position = "identity") +
  geom_text_repel(data = filter_data_LA(40)[nrow(filter_data_LA(40)), ], aes(label = round(launch_angle)), color = "red", position = "identity") + 
  geom_text_repel(data = filter_data_LA(50)[nrow(filter_data_LA(50)), ], aes(label = round(launch_angle)), color = "red", position = "identity") +
  labs(title = "Exit Velocity With Respect to Attack Angle and Centerline Angle") + 
  xlab("Centerline Angle (deg)") + ylab("Attack Angle (deg)")

```

## Plot Scaled EV Data
```{r}
plot1 <- ggplot(data_AA_CA, aes(CA, AA, z = scaled_ev)) + stat_contour(binwidth = 0.01) +
  geom_text_contour(aes(label = round(after_stat(level), 5))) + geom_abline(slope = 1, intercept = 0, color = "black", linetype = "solid") +
  geom_smooth(data = filter_data_LA(-10), se = FALSE, linetype = "dashed", color = "red") +
  geom_smooth(data = filter_data_LA(0), se = FALSE, linetype = "dashed", color = "red") +
  geom_smooth(data = filter_data_LA(10), se = FALSE, linetype = "dashed", color = "red") +
  geom_smooth(data = filter_data_LA(20), se = FALSE, linetype = "dashed", color = "red") +
  geom_smooth(data = filter_data_LA(30), se = FALSE, linetype = "dashed", color = "red") +
  geom_smooth(data = filter_data_LA(40), se = FALSE, linetype = "dashed", color = "red") +
  geom_smooth(data = filter_data_LA(50), se = FALSE, linetype = "dashed", color = "red") +
  geom_text_repel(data = filter_data_LA(0)[nrow(filter_data_LA(0)), ], aes(label = paste("LA: ", round(launch_angle))), color = "red", position = "identity") +
  geom_text_repel(data = filter_data_LA(10)[nrow(filter_data_LA(10)), ], aes(label = round(launch_angle)), color = "red", position = "identity") +
  geom_text_repel(data = filter_data_LA(20)[nrow(filter_data_LA(20)), ], aes(label = round(launch_angle)), color = "red", position = "identity") +
  geom_text_repel(data = filter_data_LA(30)[nrow(filter_data_LA(30)), ], aes(label = round(launch_angle)), color = "red", position = "identity") +
  geom_text_repel(data = filter_data_LA(40)[nrow(filter_data_LA(40)), ], aes(label = round(launch_angle)), color = "red", position = "identity") +
  geom_text_repel(data = filter_data_LA(50)[nrow(filter_data_LA(50)), ], aes(label = round(launch_angle)), color = "red", position = "identity") +
  labs(title = "Scaled Exit Velocity With Respect to Attack Angle and Centerline Angle") +
  xlab("Centerline Angle (deg)") + ylab("Attack Angle (deg)")

plot1
```
## Plotting LA vs EV by AA
```{r}
ggplot(data_AA_CA %>% filter(AA == 5)) + geom_point(aes(launch_angle, exit_speed)) + geom_point(data=data_AA_CA %>% filter(AA == 15), aes(launch_angle, exit_speed)) + geom_point(data=data_AA_CA %>% filter(AA == 25), aes(launch_angle, exit_speed))
```

# 3. Comparing Excel Data with Quadratic Fit

## Comparing Plots with Original Excel Data and Quadratic Fit
```{r}
compare_plots <- function(name, aa, lower_la=-90, upper_la=90) {
  batter <- name_converter(name)
  #Narrow data into Batter Name, Exit Velocity, and Launch Angle with non NA value
  data_batter <- sc %>%
    filter(type == "X") %>%
    select(player_name, launch_speed, launch_angle) %>%
    na.omit() %>%
    #Label data with launch angle for every bins of width 5 degrees
    filter(player_name == batter) %>%
    mutate(theta5 = 5 * round(launch_angle / 5, 0)) %>%
    arrange(launch_angle)
  
  #Create data set of Top 15 percent exit velocity within each bins
  new_df <- data.frame(theta5 = 0, top_15_ev = 0)
  
  for (theta in unique(data_batter$theta5)) {
    n_data <- data_batter %>% 
      filter(theta5 == theta)
    df <- data.frame(theta5 = theta, top_15_ev = quantile(n_data$launch_speed, seq(0, 1, 1/20))[18])
    new_df <- rbind(new_df, df)
  }

  new_df <- new_df[-c(1),]
  
  data_batter <- inner_join(data_batter, new_df, by = "theta5")
  
  top_15_data <- data_batter %>%
    filter(launch_speed >= top_15_ev,
           launch_angle >= lower_la & launch_angle <= upper_la) %>%
    mutate(top_15 = "Yes")
    
  #Join two data
  combined_data <- left_join(data_batter, top_15_data, by = c("player_name", "launch_speed", "launch_angle", "theta5")) %>%
    filter(launch_angle >= lower_la & launch_angle <= upper_la)
  
  #Indicate whether or not the exit velocity is in Top 15 percent
  combined_data <- combined_data %>%
    mutate(top_15 = ifelse(is.na(top_15), "No", top_15))
  
  #Create a quadratic regression for Top 15 percent data
  top_15_data$launch_angle2 = top_15_data$launch_angle^2
  
  quadratic_fit <- lm(launch_speed ~ launch_angle + launch_angle2, data = top_15_data)
  ev_max <- max(predict(quadratic_fit))
  top_15_data$scaled_ev = top_15_data$launch_speed / ev_max
  combined_data <- combined_data %>%
    mutate(scaled_ev = launch_speed / ev_max)
  
  #filter contour plot scaled_ev values with desired attack angle 
  data_aa <- data_AA_CA %>%
    filter(AA == aa)
  
  ggplot(combined_data) +
    geom_point(aes(launch_angle, scaled_ev, color = top_15)) +
    scale_color_manual(values = c("red", "blue"), labels=c("Top 15", "Other")) +
    stat_smooth(method = "lm", formula = y ~ x + I(x^2), data = top_15_data, aes(launch_angle, scaled_ev), color = "black", se = FALSE) +
    geom_vline(aes(xintercept = top_15_data$launch_angle[which.max(predict(quadratic_fit))]), linetype = "dashed") +
    ggtitle(paste("Exit Velocity vs Launch Angle of", name)) +
    scale_x_continuous(name = "Launch Angle (deg)", breaks = seq(-90, 90, by = 10)) +
    scale_y_continuous(name = "Scaled Exit Velocity (mph)", limits=c(0.1, 1.1), breaks = seq(0, 1.1, by = 0.1)) +
    annotate(geom = "text", label = top_15_data$launch_angle[which.max(predict(quadratic_fit))], x = top_15_data$launch_angle[which.max(predict(quadratic_fit))] + 3, y = 0.33, fontface = 'bold', vjust = 1) +
    #Combine with contour plot data
    geom_point(data = data_aa, aes(launch_angle, scaled_ev_by_AA), color="green") +
    annotate(geom = "text", label = paste("AA =", aa), x = -60, y = 1.1, color='green', fontface = 'bold', vjust = 1, size=5)

}

#Testing plots
# compare_plots("Altuve, Jose", 17, -90, 90)

```


# 4. Deep Analysis

## Comparing Plots for Various Players
```{r}
# 5 deg Quad Peak LA, Hawk AA

compare_plots("Yuli Gurriel", 7) #-2, 4

compare_plots("Yordan Alvarez", 15.5) # 17, 10

compare_plots("Freddie Freeman", 15) # 16, 10.5

compare_plots("Trea Turner", 10) # 8, 11.5

compare_plots("Kyle Tucker", 17) # 18, 12

compare_plots("Jeremy Peña", 11) # 10, 12.5

compare_plots("Will Smith", 16.5) # 18, 15

compare_plots("Alex Bregman", 11.5) # 11, 16

compare_plots("Mookie Betts", 16) # 16, 16.5

compare_plots("Jose Altuve", 17.5) # 20, 19

```
## Finding the optimal number of bin size and percentage for quadratic fit
```{r}
# 7 deg, 5 pct
peak_las <- c()
for (batter in hawk_data_inplay_player$batter_name) {
  peak_las <- c(peak_las, la_for_peak_ev(batter)[[1]])
}

hawk_data_inplay_player$peak_la_7_deg_5_pct <- peak_las

hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Diaz, Aledmys", "peak_la_7_deg_5_pct"] <- la_for_peak_ev("Díaz, Aledmys")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Dubon, Mauricio", "peak_la_7_deg_5_pct"] <- la_for_peak_ev("Dubón, Mauricio")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Maldonado, Martin", "peak_la_7_deg_5_pct"] <- la_for_peak_ev("Maldonado, Martín")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Mancini III, Trey", "peak_la_7_deg_5_pct"] <- la_for_peak_ev("Mancini, Trey")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Pena, Jeremy", "peak_la_7_deg_5_pct"] <- la_for_peak_ev("Peña, Jeremy")[[1]]

# 7 deg, 10 pct
peak_las <- c()
for (batter in hawk_data_inplay_player$batter_name) {
  peak_las <- c(peak_las, la_for_peak_ev(batter)[[1]])
}

hawk_data_inplay_player$peak_la_7_deg_10_pct <- peak_las

hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Diaz, Aledmys", "peak_la_7_deg_10_pct"] <- la_for_peak_ev("Díaz, Aledmys")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Dubon, Mauricio", "peak_la_7_deg_10_pct"] <- la_for_peak_ev("Dubón, Mauricio")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Maldonado, Martin", "peak_la_7_deg_10_pct"] <- la_for_peak_ev("Maldonado, Martín")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Mancini III, Trey", "peak_la_7_deg_10_pct"] <- la_for_peak_ev("Mancini, Trey")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Pena, Jeremy", "peak_la_7_deg_10_pct"] <- la_for_peak_ev("Peña, Jeremy")[[1]]

# 7 deg, 15 pct
peak_las <- c()
for (batter in hawk_data_inplay_player$batter_name) {
  peak_las <- c(peak_las, la_for_peak_ev(batter)[[1]])
}

hawk_data_inplay_player$peak_la_7_deg_15_pct <- peak_las

hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Diaz, Aledmys", "peak_la_7_deg_15_pct"] <- la_for_peak_ev("Díaz, Aledmys")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Dubon, Mauricio", "peak_la_7_deg_15_pct"] <- la_for_peak_ev("Dubón, Mauricio")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Maldonado, Martin", "peak_la_7_deg_15_pct"] <- la_for_peak_ev("Maldonado, Martín")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Mancini III, Trey", "peak_la_7_deg_15_pct"] <- la_for_peak_ev("Mancini, Trey")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Pena, Jeremy", "peak_la_7_deg_15_pct"] <- la_for_peak_ev("Peña, Jeremy")[[1]]

# 7 deg, 20 pct
peak_las <- c()
for (batter in hawk_data_inplay_player$batter_name) {
  peak_las <- c(peak_las, la_for_peak_ev(batter)[[1]])
}

hawk_data_inplay_player$peak_la_7_deg_20_pct <- peak_las

hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Diaz, Aledmys", "peak_la_7_deg_20_pct"] <- la_for_peak_ev("Díaz, Aledmys")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Dubon, Mauricio", "peak_la_7_deg_20_pct"] <- la_for_peak_ev("Dubón, Mauricio")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Maldonado, Martin", "peak_la_7_deg_20_pct"] <- la_for_peak_ev("Maldonado, Martín")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Mancini III, Trey", "peak_la_7_deg_20_pct"] <- la_for_peak_ev("Mancini, Trey")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Pena, Jeremy", "peak_la_7_deg_20_pct"] <- la_for_peak_ev("Peña, Jeremy")[[1]]

# 6 deg, 5 pct
peak_las <- c()
for (batter in hawk_data_inplay_player$batter_name) {
  peak_las <- c(peak_las, la_for_peak_ev(batter)[[1]])
}

hawk_data_inplay_player$peak_la_6_deg_5_pct <- peak_las

hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Diaz, Aledmys", "peak_la_6_deg_5_pct"] <- la_for_peak_ev("Díaz, Aledmys")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Dubon, Mauricio", "peak_la_6_deg_5_pct"] <- la_for_peak_ev("Dubón, Mauricio")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Maldonado, Martin", "peak_la_6_deg_5_pct"] <- la_for_peak_ev("Maldonado, Martín")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Mancini III, Trey", "peak_la_6_deg_5_pct"] <- la_for_peak_ev("Mancini, Trey")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Pena, Jeremy", "peak_la_6_deg_5_pct"] <- la_for_peak_ev("Peña, Jeremy")[[1]]

# 6 deg, 10 pct
peak_las <- c()
for (batter in hawk_data_inplay_player$batter_name) {
  peak_las <- c(peak_las, la_for_peak_ev(batter)[[1]])
}

hawk_data_inplay_player$peak_la_6_deg_10_pct <- peak_las

hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Diaz, Aledmys", "peak_la_6_deg_10_pct"] <- la_for_peak_ev("Díaz, Aledmys")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Dubon, Mauricio", "peak_la_6_deg_10_pct"] <- la_for_peak_ev("Dubón, Mauricio")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Maldonado, Martin", "peak_la_6_deg_10_pct"] <- la_for_peak_ev("Maldonado, Martín")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Mancini III, Trey", "peak_la_6_deg_10_pct"] <- la_for_peak_ev("Mancini, Trey")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Pena, Jeremy", "peak_la_6_deg_10_pct"] <- la_for_peak_ev("Peña, Jeremy")[[1]]

# 6 deg, 15 pct
peak_las <- c()
for (batter in hawk_data_inplay_player$batter_name) {
  peak_las <- c(peak_las, la_for_peak_ev(batter)[[1]])
}

hawk_data_inplay_player$peak_la_6_deg_15_pct <- peak_las

hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Diaz, Aledmys", "peak_la_6_deg_15_pct"] <- la_for_peak_ev("Díaz, Aledmys")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Dubon, Mauricio", "peak_la_6_deg_15_pct"] <- la_for_peak_ev("Dubón, Mauricio")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Maldonado, Martin", "peak_la_6_deg_15_pct"] <- la_for_peak_ev("Maldonado, Martín")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Mancini III, Trey", "peak_la_6_deg_15_pct"] <- la_for_peak_ev("Mancini, Trey")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Pena, Jeremy", "peak_la_6_deg_15_pct"] <- la_for_peak_ev("Peña, Jeremy")[[1]]

# 6 deg, 20 pct
peak_las <- c()
for (batter in hawk_data_inplay_player$batter_name) {
  peak_las <- c(peak_las, la_for_peak_ev(batter)[[1]])
}

hawk_data_inplay_player$peak_la_6_deg_20_pct <- peak_las

hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Diaz, Aledmys", "peak_la_6_deg_20_pct"] <- la_for_peak_ev("Díaz, Aledmys")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Dubon, Mauricio", "peak_la_6_deg_20_pct"] <- la_for_peak_ev("Dubón, Mauricio")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Maldonado, Martin", "peak_la_6_deg_20_pct"] <- la_for_peak_ev("Maldonado, Martín")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Mancini III, Trey", "peak_la_6_deg_20_pct"] <- la_for_peak_ev("Mancini, Trey")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Pena, Jeremy", "peak_la_6_deg_20_pct"] <- la_for_peak_ev("Peña, Jeremy")[[1]]

# 5 deg, 5 pct
peak_las <- c()
for (batter in hawk_data_inplay_player$batter_name) {
  peak_las <- c(peak_las, la_for_peak_ev(batter)[[1]])
}

hawk_data_inplay_player$peak_la_5_deg_5_pct <- peak_las

hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Diaz, Aledmys", "peak_la_5_deg_5_pct"] <- la_for_peak_ev("Díaz, Aledmys")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Dubon, Mauricio", "peak_la_5_deg_5_pct"] <- la_for_peak_ev("Dubón, Mauricio")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Maldonado, Martin", "peak_la_5_deg_5_pct"] <- la_for_peak_ev("Maldonado, Martín")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Mancini III, Trey", "peak_la_5_deg_5_pct"] <- la_for_peak_ev("Mancini, Trey")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Pena, Jeremy", "peak_la_5_deg_5_pct"] <- la_for_peak_ev("Peña, Jeremy")[[1]]

# 5 deg, 10 pct
peak_las <- c()
for (batter in hawk_data_inplay_player$batter_name) {
  peak_las <- c(peak_las, la_for_peak_ev(batter)[[1]])
}

hawk_data_inplay_player$peak_la_5_deg_10_pct <- peak_las

hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Diaz, Aledmys", "peak_la_5_deg_10_pct"] <- la_for_peak_ev("Díaz, Aledmys")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Dubon, Mauricio", "peak_la_5_deg_10_pct"] <- la_for_peak_ev("Dubón, Mauricio")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Maldonado, Martin", "peak_la_5_deg_10_pct"] <- la_for_peak_ev("Maldonado, Martín")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Mancini III, Trey", "peak_la_5_deg_10_pct"] <- la_for_peak_ev("Mancini, Trey")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Pena, Jeremy", "peak_la_5_deg_10_pct"] <- la_for_peak_ev("Peña, Jeremy")[[1]]

# 5 deg, 15 pct
peak_las <- c()
for (batter in hawk_data_inplay_player$batter_name) {
  peak_las <- c(peak_las, la_for_peak_ev(batter)[[1]])
}

hawk_data_inplay_player$peak_la_5_deg_15_pct <- peak_las

hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Diaz, Aledmys", "peak_la_5_deg_15_pct"] <- la_for_peak_ev("Díaz, Aledmys")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Dubon, Mauricio", "peak_la_5_deg_15_pct"] <- la_for_peak_ev("Dubón, Mauricio")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Maldonado, Martin", "peak_la_5_deg_15_pct"] <- la_for_peak_ev("Maldonado, Martín")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Mancini III, Trey", "peak_la_5_deg_15_pct"] <- la_for_peak_ev("Mancini, Trey")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Pena, Jeremy", "peak_la_5_deg_15_pct"] <- la_for_peak_ev("Peña, Jeremy")[[1]]

# 5 deg, 20 pct
peak_las <- c()
for (batter in hawk_data_inplay_player$batter_name) {
  peak_las <- c(peak_las, la_for_peak_ev(batter)[[1]])
}

hawk_data_inplay_player$peak_la_5_deg_20_pct <- peak_las

hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Diaz, Aledmys", "peak_la_5_deg_20_pct"] <- la_for_peak_ev("Díaz, Aledmys")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Dubon, Mauricio", "peak_la_5_deg_20_pct"] <- la_for_peak_ev("Dubón, Mauricio")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Maldonado, Martin", "peak_la_5_deg_20_pct"] <- la_for_peak_ev("Maldonado, Martín")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Mancini III, Trey", "peak_la_5_deg_20_pct"] <- la_for_peak_ev("Mancini, Trey")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Pena, Jeremy", "peak_la_5_deg_20_pct"] <- la_for_peak_ev("Peña, Jeremy")[[1]]

# 4 deg, 5 pct
peak_las <- c()
for (batter in hawk_data_inplay_player$batter_name) {
  peak_las <- c(peak_las, la_for_peak_ev(batter)[[1]])
}

hawk_data_inplay_player$peak_la_4_deg_5_pct <- peak_las

hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Diaz, Aledmys", "peak_la_4_deg_5_pct"] <- la_for_peak_ev("Díaz, Aledmys")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Dubon, Mauricio", "peak_la_4_deg_5_pct"] <- la_for_peak_ev("Dubón, Mauricio")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Maldonado, Martin", "peak_la_4_deg_5_pct"] <- la_for_peak_ev("Maldonado, Martín")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Mancini III, Trey", "peak_la_4_deg_5_pct"] <- la_for_peak_ev("Mancini, Trey")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Pena, Jeremy", "peak_la_4_deg_5_pct"] <- la_for_peak_ev("Peña, Jeremy")[[1]]

# 4 deg, 10 pct
peak_las <- c()
for (batter in hawk_data_inplay_player$batter_name) {
  peak_las <- c(peak_las, la_for_peak_ev(batter)[[1]])
}

hawk_data_inplay_player$peak_la_4_deg_10_pct <- peak_las

hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Diaz, Aledmys", "peak_la_4_deg_10_pct"] <- la_for_peak_ev("Díaz, Aledmys")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Dubon, Mauricio", "peak_la_4_deg_10_pct"] <- la_for_peak_ev("Dubón, Mauricio")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Maldonado, Martin", "peak_la_4_deg_10_pct"] <- la_for_peak_ev("Maldonado, Martín")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Mancini III, Trey", "peak_la_4_deg_10_pct"] <- la_for_peak_ev("Mancini, Trey")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Pena, Jeremy", "peak_la_4_deg_10_pct"] <- la_for_peak_ev("Peña, Jeremy")[[1]]

# 4 deg, 15 pct
peak_las <- c()
for (batter in hawk_data_inplay_player$batter_name) {
  peak_las <- c(peak_las, la_for_peak_ev(batter)[[1]])
}

hawk_data_inplay_player$peak_la_4_deg_15_pct <- peak_las

hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Diaz, Aledmys", "peak_la_4_deg_15_pct"] <- la_for_peak_ev("Díaz, Aledmys")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Dubon, Mauricio", "peak_la_4_deg_15_pct"] <- la_for_peak_ev("Dubón, Mauricio")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Maldonado, Martin", "peak_la_4_deg_15_pct"] <- la_for_peak_ev("Maldonado, Martín")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Mancini III, Trey", "peak_la_4_deg_15_pct"] <- la_for_peak_ev("Mancini, Trey")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Pena, Jeremy", "peak_la_4_deg_15_pct"] <- la_for_peak_ev("Peña, Jeremy")[[1]]

# 4 deg, 20 pct
peak_las <- c()
for (batter in hawk_data_inplay_player$batter_name) {
  peak_las <- c(peak_las, la_for_peak_ev(batter)[[1]])
}

hawk_data_inplay_player$peak_la_5_deg_20_pct <- peak_las

hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Diaz, Aledmys", "peak_la_4_deg_20_pct"] <- la_for_peak_ev("Díaz, Aledmys")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Dubon, Mauricio", "peak_la_4_deg_20_pct"] <- la_for_peak_ev("Dubón, Mauricio")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Maldonado, Martin", "peak_la_4_deg_20_pct"] <- la_for_peak_ev("Maldonado, Martín")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Mancini III, Trey", "peak_la_4_deg_20_pct"] <- la_for_peak_ev("Mancini, Trey")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Pena, Jeremy", "peak_la_4_deg_20_pct"] <- la_for_peak_ev("Peña, Jeremy")[[1]]

# 3 deg 5 pct
peak_las <- c()
for (batter in hawk_data_inplay_player$batter_name) {
  peak_las <- c(peak_las, la_for_peak_ev(batter)[[1]])
}

hawk_data_inplay_player$peak_la_3_deg_5_pct <- peak_las

hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Diaz, Aledmys", "peak_la_3_deg_5_pct"] <- la_for_peak_ev("Díaz, Aledmys")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Dubon, Mauricio", "peak_la_3_deg_5_pct"] <- la_for_peak_ev("Dubón, Mauricio")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Maldonado, Martin", "peak_la_3_deg_5_pct"] <- la_for_peak_ev("Maldonado, Martín")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Mancini III, Trey", "peak_la_3_deg_5_pct"] <- la_for_peak_ev("Mancini, Trey")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Pena, Jeremy", "peak_la_3_deg_5_pct"] <- la_for_peak_ev("Peña, Jeremy")[[1]]

# 3 deg 10 pct
peak_las <- c()
for (batter in hawk_data_inplay_player$batter_name) {
  peak_las <- c(peak_las, la_for_peak_ev(batter)[[1]])
}

hawk_data_inplay_player$peak_la_3_deg_10_pct <- peak_las

hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Diaz, Aledmys", "peak_la_3_deg_10_pct"] <- la_for_peak_ev("Díaz, Aledmys")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Dubon, Mauricio", "peak_la_3_deg_10_pct"] <- la_for_peak_ev("Dubón, Mauricio")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Maldonado, Martin", "peak_la_3_deg_10_pct"] <- la_for_peak_ev("Maldonado, Martín")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Mancini III, Trey", "peak_la_3_deg_10_pct"] <- la_for_peak_ev("Mancini, Trey")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Pena, Jeremy", "peak_la_3_deg_10_pct"] <- la_for_peak_ev("Peña, Jeremy")[[1]]

# 3 deg 15 pct
peak_las <- c()
for (batter in hawk_data_inplay_player$batter_name) {
  peak_las <- c(peak_las, la_for_peak_ev(batter)[[1]])
}

hawk_data_inplay_player$peak_la_3_deg_15_pct <- peak_las

hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Diaz, Aledmys", "peak_la_3_deg_15_pct"] <- la_for_peak_ev("Díaz, Aledmys")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Dubon, Mauricio", "peak_la_3_deg_15_pct"] <- la_for_peak_ev("Dubón, Mauricio")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Maldonado, Martin", "peak_la_3_deg_15_pct"] <- la_for_peak_ev("Maldonado, Martín")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Mancini III, Trey", "peak_la_3_deg_15_pct"] <- la_for_peak_ev("Mancini, Trey")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Pena, Jeremy", "peak_la_3_deg_15_pct"] <- la_for_peak_ev("Peña, Jeremy")[[1]]

```

```{r}
model_7_deg_5_pct <- lm(peak_la_7_deg_5_pct ~ median_attack_angle, data=hawk_data_inplay_player)
summary(model_7_deg_5_pct)$r.squared

model_7_deg_10_pct <- lm(peak_la_7_deg_10_pct ~ median_attack_angle, data=hawk_data_inplay_player)
summary(model_7_deg_10_pct)$r.squared

model_7_deg_15_pct <- lm(peak_la_7_deg_15_pct ~ median_attack_angle, data=hawk_data_inplay_player)
summary(model_7_deg_15_pct)$r.squared

model_7_deg_20_pct <- lm(peak_la_7_deg_20_pct ~ median_attack_angle, data=hawk_data_inplay_player)
summary(model_7_deg_20_pct)$r.squared

model_6_deg_5_pct <- lm(peak_la_6_deg_5_pct ~ median_attack_angle, data=hawk_data_inplay_player)
summary(model_6_deg_5_pct)$r.squared

model_6_deg_10_pct <- lm(peak_la_6_deg_10_pct ~ median_attack_angle, data=hawk_data_inplay_player)
summary(model_6_deg_10_pct)$r.squared

model_6_deg_15_pct <- lm(peak_la_6_deg_15_pct ~ median_attack_angle, data=hawk_data_inplay_player)
summary(model_6_deg_15_pct)$r.squared

model_6_deg_20_pct <- lm(peak_la_6_deg_20_pct ~ median_attack_angle, data=hawk_data_inplay_player)
summary(model_6_deg_20_pct)$r.squared

model_5_deg_5_pct <- lm(peak_la_5_deg_5_pct ~ median_attack_angle, data=hawk_data_inplay_player)
summary(model_5_deg_5_pct)$r.squared

model_5_deg_10_pct <- lm(peak_la_5_deg_10_pct ~ median_attack_angle, data=hawk_data_inplay_player)
summary(model_5_deg_10_pct)$r.squared

model_5_deg_15_pct <- lm(peak_la_5_deg_15_pct ~ median_attack_angle, data=hawk_data_inplay_player)
summary(model_5_deg_15_pct)$r.squared

model_5_deg_20_pct <- lm(peak_la_5_deg_20_pct ~ median_attack_angle, data=hawk_data_inplay_player)
summary(model_5_deg_20_pct)$r.squared # 0.617467

model_4_deg_5_pct <- lm(peak_la_4_deg_5_pct ~ median_attack_angle, data=hawk_data_inplay_player)
summary(model_4_deg_5_pct)$r.squared

model_4_deg_10_pct <- lm(peak_la_4_deg_10_pct ~ median_attack_angle, data=hawk_data_inplay_player)
summary(model_4_deg_10_pct)$r.squared

model_4_deg_15_pct <- lm(peak_la_4_deg_15_pct ~ median_attack_angle, data=hawk_data_inplay_player)
summary(model_4_deg_15_pct)$r.squared

model_4_deg_20_pct <- lm(peak_la_4_deg_20_pct ~ median_attack_angle, data=hawk_data_inplay_player)
summary(model_4_deg_20_pct)$r.squared

model_3_deg_5_pct <- lm(peak_la_3_deg_5_pct ~ median_attack_angle, data=hawk_data_inplay_player)
summary(model_3_deg_5_pct)$r.squared

model_3_deg_10_pct <- lm(peak_la_3_deg_10_pct ~ median_attack_angle, data=hawk_data_inplay_player)
summary(model_3_deg_10_pct)$r.squared

model_3_deg_15_pct <- lm(peak_la_3_deg_15_pct ~ median_attack_angle, data=hawk_data_inplay_player)
summary(model_3_deg_15_pct)$r.squared
```

## Compare Peak Launch Angle with AA from Hawkeye
```{r}
# 5 deg, 15 pct
peak_las <- c()
for (batter in hawk_data_inplay_player$batter_name) {
  peak_las <- c(peak_las, la_for_peak_ev(batter)[[1]])
}

hawk_data_inplay_player$peak_la_5_deg_15_pct <- peak_las

hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Diaz, Aledmys", "peak_la_5_deg_15_pct"] <- la_for_peak_ev("Díaz, Aledmys")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Dubon, Mauricio", "peak_la_5_deg_15_pct"] <- la_for_peak_ev("Dubón, Mauricio")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Maldonado, Martin", "peak_la_5_deg_15_pct"] <- la_for_peak_ev("Maldonado, Martín")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Mancini III, Trey", "peak_la_5_deg_15_pct"] <- la_for_peak_ev("Mancini, Trey")[[1]]
hawk_data_inplay_player[hawk_data_inplay_player$batter_name == "Pena, Jeremy", "peak_la_5_deg_15_pct"] <- la_for_peak_ev("Peña, Jeremy")[[1]]

# Plot 
ggplot(hawk_data_inplay_player) + geom_point(aes(median_attack_angle, peak_la_5_deg_15_pct)) + geom_smooth(data=data_AA_CA_max_ev, aes(AA, launch_angle), method='lm', color='blue') + xlim(0, 25) + ylim(-20, 40) + labs(y="Peak Launch Angle (deg)", x="Median Attack Angle (deg)") + annotate(geom="text", x=15, y=-5, label="Simple Ball-Bat Collision Model", color='blue') #+ theme(axis.text = element_text(size = 14), axis.title = element_text(size = 16), plot.title = element_text(size = 18))
```




# To do:
- Find relationship between Peak LA and AA.







# Not good method below.
## Plot each player's data on AA vs CA graph
```{r}
get_player_data <- function(player_name) {
  data_player <- data_AA_CA %>%
    filter(round(launch_angle) == round(as.numeric(la_for_peak_ev(player_name)[1])),
         CA == AA) %>%
    mutate(player_name = player_name)
  return(data_player[1,])
}

top_hr_hitters <- c("Perez, Salvador", "Guerrero Jr., Vladimir", "Ohtani, Shohei", "Semien, Marcus", "Tatis Jr., Fernando", "Judge, Aaron", "Olson, Matt", "Haniger, Mitch" ,"Lowe, Brandon", "Devers, Rafael")

top_avg_hitters <- c("Turner, Trea", "Gurriel, Yuli", "Soto, Juan", "Brantley, Michael", "Guerrero Jr., Vladimir", "Marte, Starling", "Harper, Bryce", "Anderson, Tim", "Castellanos, Nick", "Frazier, Adam") 


get_player_df <- function(players) {
  data_players <- get_player_data(players[1])
  for (i in 2: length(players)) {
  new_data_player <- get_player_data((players[i]))
  data_players <- rbind(data_players, new_data_player)
  }
  return(data_players)
}

plot1 + geom_point(data = get_player_df(top_hr_hitters), aes(CA, AA)) + geom_text_repel(data = get_player_df(top_hr_hitters), aes(label = player_name))

plot1 + geom_point(data = get_player_df(top_avg_hitters), aes(CA, AA)) + geom_text_repel(data = get_player_df(top_avg_hitters), aes(label = player_name))
```




## Plot each player's Top 15 exit velo on AA vs CA graph
```{r}
#Function for plotting Exit Velocity vs Launch Angle for a batter
get_top_15_data <- function(batter) {
  #Narrow data into Batter Name, Exit Velocity, and Launch Angle with non NA value
  data_batter <- new_data2 %>%
  select(player_name, scaled_ev, launch_speed, launch_angle) %>%
  na.omit() %>%
  #Label data with launch angle for every bins of width 5 degrees
  filter(player_name == batter) %>%
  mutate(theta5 = 5 * round(launch_angle / 5, 0)) %>%
  arrange(launch_angle)
  
  #Create data set of Top 15 percent exit velocity within each bins
  new_df <- data.frame(theta5 = 0, top_15_ev = 0)
  
  for (theta in unique(data_batter$theta5)) {
    n_data <- data_batter %>% 
      filter(theta5 == theta)
    df <- data.frame(theta5 = theta, top_15_ev = quantile(n_data$scaled_ev, seq(0, 1, 1/10))[10])
    new_df <- rbind(new_df, df)
  }

  new_df <- new_df[-c(1),]
  
  data_batter <- inner_join(data_batter, new_df, by = "theta5")
  
  top_15_data <- data_batter %>%
    filter(scaled_ev >= top_15_ev,
           launch_angle > -40) %>%
    mutate(top_15 = "Yes")
  return(top_15_data)
}
  
#Get closest data in excel

get_approx_data <- function(player) {
  player_data <- get_top_15_data(player) %>%
  filter(launch_angle >= min(data_AA_CA$launch_angle),
         launch_angle <= max(data_AA_CA$launch_angle))
  

  data_o <- data_AA_CA %>%
  filter(launch_angle < 1 + player_data[1, ]$launch_angle,
         launch_angle > -1 + player_data[1, ]$launch_angle)
  data_oo <- data_o[which.min(abs(as.numeric(player_data[1, ]$scaled_ev)-data_o$scaled_ev)),]
  
  for (i in 2:nrow(player_data)) {
    new_data_o <- data_AA_CA %>%
      filter(launch_angle < 1 + player_data[i, ]$launch_angle,
             launch_angle > -1 + player_data[i, ]$launch_angle)
    new_data_oo <- new_data_o[which.min(abs(as.numeric(player_data[i, ]$scaled_ev)-new_data_o$scaled_ev)),]
    data_oo <- rbind(data_oo, new_data_oo)
  }
  
  return(data_oo)
}

plot1 + geom_point(data = get_approx_data("Ohtani, Shohei"), aes(CA, AA))
```

```{r}
#Function for plotting Exit Velocity vs Launch Angle for a batter
get_all_data <- function(batter) {
  data_batter <- new_data2 %>%
  #Label data with launch angle for every bins of width 5 degrees
  filter(player_name == batter) %>%
  mutate(theta5 = 5 * round(launch_angle / 5, 0)) %>%
  arrange(launch_angle)
  
  #Create data set of Top 15 percent exit velocity within each bins
  new_df <- data.frame(theta5 = 0, top_15_ev = 0)
  
  for (theta in unique(data_batter$theta5)) {
    n_data <- data_batter %>% 
      filter(theta5 == theta)
    df <- data.frame(theta5 = theta, top_15_ev = quantile(n_data$scaled_ev, seq(0, 1, 1/10))[10])
    new_df <- rbind(new_df, df)
  }

  new_df <- new_df[-c(1),]
  
  data_batter <- inner_join(data_batter, new_df, by = "theta5")
  
  top_15_data <- data_batter %>%
    filter(scaled_ev >= top_15_ev,
           launch_angle > -40) %>%
    mutate(top_15 = "Yes")
  
  combined_data <- left_join(data_batter, top_15_data, by = c("player_name", "scaled_ev", "launch_speed", "launch_angle", "theta5", "description", "events", "top_15_ev", "max_ev"))
    #Indicate whether or not the exit velocity is in Top 15 percent
  combined_data <- combined_data %>%
    mutate(top_15 = ifelse(is.na(top_15), "No", top_15)) %>%
    filter(launch_angle > -40) %>%
    mutate(events = ifelse(events == "", "foul", events))
  
  #Reduce number of different types of events, excluding bunts, catcher interf, game advisory
  reduced_events_data <- combined_data %>%
    mutate(reduced_events = case_when(events %in% c("single") ~ "single",
                                      events %in% c("double", "triple") ~ "double & triple",
                                      events %in% c("home_run") ~ "homerun",
                                      events %in% c("field_out", "grounded_into_double_play", "field_error", "force_out", "sac_fly", "fielders_choice", "double_play", "fielders_choice_out", "sac_fly_double_play", "triple_play") ~ "out",
                                      events == "foul" ~ "foul")) %>%
    na.omit()

  return(reduced_events_data)
}
  
#Get closest data in excel

get_approx_data2 <- function(player) {
  player_data <- get_all_data(player) %>%
  filter(launch_angle >= min(data_AA_CA$launch_angle),
         launch_angle <= max(data_AA_CA$launch_angle))
  

  data_o <- data_AA_CA %>%
  filter(launch_angle < 1 + player_data[1, ]$launch_angle,
         launch_angle > -1 + player_data[1, ]$launch_angle)
  data_oo <- data_o[which.min(abs(as.numeric(player_data[1, ]$scaled_ev)-data_o$scaled_ev)),]
  
  for (i in 2:nrow(player_data)) {
    new_data_o <- data_AA_CA %>%
      filter(launch_angle < 1 + player_data[i, ]$launch_angle,
             launch_angle > -1 + player_data[i, ]$launch_angle)
    new_data_oo <- new_data_o[which.min(abs(as.numeric(player_data[i, ]$scaled_ev)-new_data_o$scaled_ev)),]
    data_oo <- rbind(data_oo, new_data_oo)
  }
  return(data_oo)
}

#######################################################################################################
plot1 + geom_point(data = get_approx_data2("Ohtani, Shohei"), aes(CA, AA))

ggplot(get_approx_data2("Altuve, Jose")) + geom_point(aes(CA, AA))

  do2 <- get_all_data("Ohtani, Shohei") %>%
  filter(launch_angle >= min(data_AA_CA$launch_angle),
         launch_angle <= max(data_AA_CA$launch_angle))
  

  do3 <- data_AA_CA %>%
  filter(launch_angle < 1 + do2[1, ]$launch_angle,
         launch_angle > -1 + do2[1, ]$launch_angle)
  do4 <- do3[which.min(abs(as.numeric(do2[1, ]$scaled_ev)-do3$scaled_ev)),]
  
  for (i in 2:nrow(do2)) {
    o2 <- data_AA_CA %>%
      filter(launch_angle < 1 + do2[i, ]$launch_angle,
             launch_angle > -1 + do2[i, ]$launch_angle)
    o3 <- o2[which.min(abs(as.numeric(do2[i, ]$scaled_ev)-o2$scaled_ev)),]
    do4 <- rbind(do4, o3)
  }
  do4 <- do4 %>%
    filter(CA >= -10 & CA <= 50,
           AA >= 00 & AA <= 40)


```

```{r}

plot1 + geom_point(data = get_approx_data("Altuve, Jose"), aes(CA, AA))

```

```{r}
AA_LA <- data_AA_CA %>%
  filter(AA == CA) %>%
  select(AA, launch_angle)

ggplot(AA_LA, aes(AA, launch_angle)) + geom_point() + geom_smooth()
```


```{r}
plot11 <- ggplot(data_AA_CA, aes(launch_angle , exit_speed, z = AA)) + geom_contour(binwidth = 0.01) 

+
  
  geom_text_contour(aes(label = round(..level.., 2))) +
  geom_abline(CA = AA) + 
  geom_smooth(data = filter_data_LA(0), se = FALSE, linetype = "dashed", color = "red") + 
  geom_smooth(data = filter_data_LA(10), se = FALSE, linetype = "dashed", color = "red") + 
  geom_smooth(data = filter_data_LA(20), se = FALSE, linetype = "dashed", color = "red") + 
  geom_smooth(data = filter_data_LA(30), se = FALSE, linetype = "dashed", color = "red") + 
  geom_smooth(data = filter_data_LA(40), se = FALSE, linetype = "dashed", color = "red") + 
  geom_smooth(data = filter_data_LA(50), se = FALSE, linetype = "dashed", color = "red") + 
  geom_text_repel(data = filter_data_LA(0)[nrow(filter_data_LA(0)), ], aes(label = paste("LA: ", round(launch_angle))), color = "red", position = "identity") + 
  geom_text_repel(data = filter_data_LA(10)[nrow(filter_data_LA(10)), ], aes(label = round(launch_angle)), color = "red", position = "identity") +
  geom_text_repel(data = filter_data_LA(20)[nrow(filter_data_LA(20)), ], aes(label = round(launch_angle)), color = "red", position = "identity") + 
  geom_text_repel(data = filter_data_LA(30)[nrow(filter_data_LA(30)), ], aes(label = round(launch_angle)), color = "red", position = "identity") +
  geom_text_repel(data = filter_data_LA(40)[nrow(filter_data_LA(40)), ], aes(label = round(launch_angle)), color = "red", position = "identity") + 
  geom_text_repel(data = filter_data_LA(50)[nrow(filter_data_LA(50)), ], aes(label = round(launch_angle)), color = "red", position = "identity") +
  labs(title = "Exit Velocity With Respect to Attack Angle and Centerline Angle") + 
  xlab("Centerline Angle (deg)") + ylab("Attack Angle (deg)")

```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
